Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-03-14 paper.tex
code

In previous section, we presented a refunctionalized AAM that in extended continuation-passing style. 
To obtain a definitional abstract interpreter, we need to transform it further to direct-style. 
At this point, we have two choices: 
1) By monadifying the continuations into monads, we obtain the \citeauthor{darais2017abstracting}'s 
abstracting definitional interpreter in monadic style.
2) By representing the extended continuation-passing style with delimited continuations,
we derive a new form of abstract interpreters in direct-style.
This coincidence should not be a surprise since there are extraordinarily rich literatures 
show the correspondence between monads and continuations as of their very beginning era
(TODO: cite Danvy/abs control/esce of fp).
In this section, we present both of them.

@@

collecting semantics and non-determinism.

TODO: use delimited continuation?

>>>>

In previous section, we presented a refunctionalized AAM that in extended continuation-passing style. 
To obtain a definitional abstract interpreter, we need to transform it further to direct-style. 
At this point, we have two choices: 
1) By monadifying the continuations into monads, we obtain the \citeauthor{darais2017abstracting}'s 
abstracting definitional interpreter in monadic style.
2) By representing the extended continuation-passing style with delimited continuations,
we derive a new form of abstract interpreters in direct-style.
This coincidence should not be a surprise since there are extraordinarily rich literatures 
show the correspondence between monads and continuations as of their very beginning era
(TODO: cite Danvy/abs control/esce of fp).
In fact, we can also just use the same caching algorithm but with side effect such as 
assignments and mutations to achieve the same definitional interpreter with pushdown
control flows.
In this section, we present both of versions using monads and delimited controls.

@@

TODO: This version does not use delimited controls.

\begin{lstlisting}
def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): (Set[VS], Cache) = {
  val config = Config(e, env, time)
  if (cache.outContains(config)) return (cache.outGet(config), cache)

  val new_cache = cache.outUpdate(config, cache.inGet(config))
  val new_time = (e::time).take(k)

  e match {
    case Let(x, App(f, ae), e) =>
      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
      val app_vss_cache = for (Clos(Lam(v, body), c_env) <- closures) yield {
        val baddr = allocBind(v, new_time)
        val new_env = c_env + (v -> baddr)
        val new_store = store.update(baddr, aeval(ae, env, store))
        aval(body, new_env, new_store, new_time, new_cache) //TODO: Do we have to use the latest cache?
      }

      val app_vss = app_vss_cache.map(_._1).foldLeft(Set[VS]())(_ ++ _)
      val app_cache = app_vss_cache.map(_._2).foldLeft(mtCache)(_.outJoin(_))

      val e_vss = for (VS(vals, time, store) <- app_vss.toList) yield {
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        aval(e, new_env, new_store, time, app_cache) //TODO: Do we have to use the latest cache?
      }
      val result = e_vss.map(_._1).foldLeft(Set[VS]())(_ ++ _)
      (result, e_vss.map(_._2).foldLeft(mtCache)(_.outJoin(_)).outUpdate(config, result))

    case ae if isAtomic(ae) =>
      val ds = aeval(ae, env, store)
      (Set(VS(ds, new_time, store)), new_cache.outUpdate(config, VS(ds, new_time, store)))
  }
}

def analyze(e: Expr) = {
  def iter(cache: Cache): (Set[VS], Cache) = {
    val (vss, new_cache) = aval(e, mtEnv, mtStore, mtTime, cache)
    if (new_cache.out == cache.out) (vss, new_cache)
    else iter (Cache(new_cache.out, new_cache.out))
  }
  iter(mtCache)
}
\end{lstlisting}
